### 2024-06-03 thesis/input_files/energymarket.tex
final

This chapter has two intents. Firstly, it summarizes several concepts for the newcomers to the field of electricity markets. Secondly, it discusses trends and challenges in the power sector. Such developments motivate the need for robust and efficient forecasting techniques.

@@

\section{Electricity markets}
Electricity markets have undergone rapid changes all over the world during the last three decades.
Before this revolution, the power sector was a natural monopoly. In particular, it was characterized by a high vertical integration and by very little market competition. Better technologies in transmission, generation and distribution are the reasons for such liberal shift. The rationale is simple, a competitive market promotes efficiency and stimulates technical innovation more than a monopoly does.
\\
During the 1990s, various wholesale electricity markets were established; for example in Chile, Great Britain, the Scandinivian block, Australia, New Zealand, the US and Canada. New participants, each with a specific task, entered the power market scenario.
\begin{itemize}
    \item Electricity generators produce electricity to be consumed.
    \item Electricity transmission owners 
    move high voltage electricity to electric utility companies whilst ensuring safe and reliable supply.
    \item Electricity grid operators are responsible for scheduling electricity over the transmission network in order to ensure supply demand balance.
    \item Electric utilities deliver power over local lines to consumers. 
    \item Retail energy suppliers purchase electricity in the wholesale market from electricity generators and resell it to consumers.
\end{itemize}
These functions are not mutually exclusive. That is, an entity can provide one or more of these services.

@@

\subsection{Price peaks}
Electricity markets are characterized by spikes in their spot prices. When this happens, the system price jumps abruptly and then drops back within a very short period.
This spikeness follows from the non storability of electricity, that is, electricity has to be consumed as it is produced. Therefore, extreme load fluctuations combined with generation outages or transmission failures can result in price spikes.

@@

\subsection{Nodal and zonal pricing}
Grid configuration and physical limits on electricity lines may lead to congestion. Zonal market clearing pricing (ZMCP) and local marginal price (LMP) are two types of pricing schemes utilized to address this. The former is adopted in European contries while the latter is used in the United States, Figure \ref{fig:eu_zonal} and Figure \ref{fig:us_local}.
With zonal market clearing pricing, prices may differ between zones but are the same within the same area.
On the other hand, local marginal price is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses. Buses is where a electricity line or several lines are connected.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/eu_zonal.png}
    \caption[Zonal pricing in the EU]{Zonal pricing in the EU \citeW{eu_zonal}}
    \label{fig:eu_zonal}
\end{figure}

@@

% pensodino \subsection{Nodal and Zonal Pricing}
\section{Smart grids}
A smart meter is a digital device recording energy consumption in real time and communicating this data to the supplier and the consumer.
Deployment of smart grids and the digitalisation of the electricity network are intended to increase the energy efficiency of our networks.
The reason is that smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
All of this motivates the need for efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector, see \cite{declean,andreadou2016telecommunication}.

\section{Renewables}
% https://www.pnnl.gov/explainer-articles/renewable-integration
A pressing challenge facing policymakers today is the energy transition towards cleaner energy, see Figure \ref{fig:electricity_production_by_source} for an up to date breakdown on energy production by source. Such move is intended to mitigate the effects of climate change and reduce pollution at the same time.
This transition passes through renewables integration into the electric grid. 
Nevertheless, integrating renewables into the existing electric grid involves several technical challenges. First, we need to develop infrastructures and technologies capable of connecting renewable power plants with the existing grid. For example, connecting renewables power plants located in remote and offshore locations to high voltage powerlines is not trivial.
Additionally, renewables such as wind and solar are characterized by seasonalities and depend heavily on weather conditions. Therefore, balancing the electricity network and mantaining stability will become harder for grid operators. They will need more flexibility in the grid and develop new approaches in order to achieve their goals.

>>>>

This chapter has two intents. Firstly, it summarises several concepts for the newcomers to the field of electricity markets. Secondly, it discusses trends and challenges in the power sector. Such developments motivate the need for robust and efficient forecasting techniques.

@@

\section{Electricity markets}
Electricity markets have undergone rapid changes all over the world during the last three decades.
Before this revolution, the power sector was a natural monopoly. In particular, it was characterised by a high vertical integration and by very little market competition. Better technologies in transmission, generation and distribution are the reasons for such liberal shift. The rationale is simple, a competitive market promotes efficiency and stimulates technical innovation more than a monopoly does.
\\
During the 1990s, various wholesale electricity markets were established; for example in Chile, Great Britain, the Scandinivian block, Australia, New Zealand, the US and Canada. New participants, each with a specific task, entered the power market scenario.
\begin{itemize}
    \item Electricity generators produce electricity to be consumed.
    \item Electricity transmission owners 
    move high voltage electricity to electric utility companies whilst ensuring safe and reliable supply.
    \item Electricity grid operators are responsible for scheduling electricity over the transmission network in order to ensure supply demand balance.
    \item Electric utilities deliver power over local lines to consumers. 
    \item Retail energy suppliers purchase electricity in the wholesale market from electricity generators and resell it to consumers.
\end{itemize}
These functions are not mutually exclusive. That is, an entity can provide one or more of these services.

@@

\subsection{Price peaks}
Electricity markets are characterised by spikes in their spot prices. When this happens, the system price jumps abruptly and then drops back within a very short period.
This spikeness follows from the non storability of electricity, that is, electricity has to be consumed as it is produced. Therefore, extreme load fluctuations combined with generation outages or transmission failures can result in price spikes.

@@

\subsection{Nodal and zonal pricing}
Grid configuration and physical limits on electricity lines may lead to congestion. Zonal market clearing pricing (ZMCP) and local marginal price (LMP) are two types of pricing schemes utilised to address this. The former is adopted in European contries while the latter is used in the United States, Figure \ref{fig:eu_zonal} and Figure \ref{fig:us_local}.
With zonal market clearing pricing, prices may differ between zones but are the same within the same area.
On the other hand, local marginal price is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses. Buses is where a electricity line or several lines are connected.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/eu_zonal.png}
    \caption[Zonal pricing in the EU]{Zonal pricing in the EU \citeW{eu_zonal}}
    \label{fig:eu_zonal}
\end{figure}

@@

% pensodino \subsection{Nodal and Zonal Pricing}
\section{Smart grids}
A smart meter is a digital device recording energy consumption in real time and communicating this data to the supplier and the consumer.
Deployment of smart grids and the digitalisation of the electricity network are intended to increase the energy efficiency of our networks.
The reason is that smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
All of this motivates the need for efficient and robust tools to analyse the vast data resulting from the digitalisation of the power sector, see \cite{declean,andreadou2016telecommunication}.

\section{Renewables}
% https://www.pnnl.gov/explainer-articles/renewable-integration
A pressing challenge facing policymakers today is the energy transition towards cleaner energy, see Figure \ref{fig:electricity_production_by_source} for an up to date breakdown on energy production by source. Such move is intended to mitigate the effects of climate change and reduce pollution at the same time.
This transition passes through renewables integration into the electric grid. 
Nevertheless, integrating renewables into the existing electric grid involves several technical challenges. First, we need to develop infrastructures and technologies capable of connecting renewable power plants with the existing grid. For example, connecting renewables power plants located in remote and offshore locations to high voltage powerlines is not trivial.
Additionally, renewables such as wind and solar are characterised by seasonalities and depend heavily on weather conditions. Therefore, balancing the electricity network and mantaining stability will become harder for grid operators. They will need more flexibility in the grid and develop new approaches in order to achieve their goals.

### 2024-06-03 thesis/input_files/implementation.tex
final

This section is intended to explain and aid for reproducibility studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
\\
For the list of Python packages needed, see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
% - indicate computer specifics
All experiments have been carried out on a 3.2 GHz 16GB Apple M1 Pro.

@@

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented Python open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point method. This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the $Q$ matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a single matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with in built scikit-learn methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}. Functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels. 
Kernels from these two submodules use a different convention in the hyperparameters. The former has the lenghtscale parameter in the numerator while the latter in the denominator. In order to meaningfully compare across different kernels we made them consistent in this aspect. For this purpose we adhere to the convention of having lenghtscale hyperparameters in the denominator.

>>>>

This section is intended to explain and aid for reproducibility studies. Hereafter, the specific libraries used and the custom implementations are thoroughly documented.
\\
For the list of Python packages needed, see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
% - indicate computer specifics
All experiments have been carried out on a 3.2 GHz 16GB Apple M1 Pro.

@@

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented Python open source library, thus, the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialised by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point algorithm. This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus, we just have to provide the $Q$ matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a single matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with in built scikit-learn methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}. Functions supported by our kernel quantile regression include: Gaussian RBF, Absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels. 
Kernels from these two submodules use a different convention in the hyperparameters. The former has the lengthscale parameter in the numerator while the latter in the denominator. In order to meaningfully compare across different kernels we made them consistent in this aspect. For this purpose we adhere to the convention of having lengthscale hyperparameters in the denominator.

### 2024-06-03 thesis/input_files/literaturereview.tex
final

%Electricity Forecasting 
%intro
During the past 25 years %because paper of 2014 says 15, so now it is 25
a wide range of new ideas have been proposed for electricity point forecasting and for probabilistic forecasting.
\\
The field benefitted greatly from the increse of computing power, the greater availability of data and the interest in data science.
As a consequence, the forecaster's toolbox has grown in size and complexity.
\\
Before delving into the literature review, we stress that at this point in time there is no superior method. Different solutions may outperform or underperform compared to other techniques depending on the problem settings. Thus, understanding the complexity, strenghts and weaknesses of each method is crucial for fitting the right model to the right setting.
\\
Within this research community, the need for more homogeneity in the choice of the error valuation metrics emerged (Section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}. As a solution, \cite{lago} proposes a checklist to aid evaluating the meaningfulness of new research. Throughout this thesis work, we will stick to the proposed principles and best practices peculiar of the EF field.
%literature review

@@

In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF). Consensus in the literature is to use one day, two weeks and three years respectilvely as cut off horizons \cite{hong_phd}; see Figure \ref{fig:time} for a visualisation. 

@@

EF literature distinguishes between point and probabilistic forecasts. Each of the two has its advantages and disadvantages. Point forecasts are easier to generate and less computationally intensive while probabilistic forecasts are more informative. Industry and research efforts have focused primarily on point forcasting. Nevertheless, interest in probabilistic forecasting has risen considerably over the last years due to renewable integration requirements, introduction of smart grids and increased market competitiveness.

@@

This section presents the results of the bibliometric analysis we performed on March $6^th$ 2024. This survey has been carried out by using the Scopus citation database. For details on the specific queries entered in Scopus, refer to Appendix \ref{src_code}.

@@

To get started, let us consider the evolution of the EF field over the years. This is visualized in Figure \ref{fig:epf_evolution}, with results grouped by category. Articles prior to the 2000 have been aggregated together due to their small number. Figure \ref{fig:epf_evolution} shows the trend of an increasing interest in EF.

@@

The next question is to compare the state of point-versus probabilistic forecasting, this is visualized in Figure \ref{fig:point_vs_prob}. What can be concluded is that probabilistic-is less developed than point forecasting. To our mind this is due to the complexity of probabilistic forecasts.

@@

EF is a heterogenous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked. Figure \ref{fig:subject_area} depicts the EF publications by subject area. What can be concluded is that the bulk of publications come from engineering, computer science, mathematics and econometrics.

@@

  \caption{Most popolar sources/outlets}

@@

Weron \cite{EPF_review} reviews the state of the art for electricity price forecasting. Besides analysing complexity of available solutions, strenghts and weaknesses, it also stresses the need for objective comparative EPF studies. Specifically, it advocates for studies using similar datasets, using the same error evaluation metric and statistical testing model's outperformance.
Hong et al.\ \cite{prob_elf} discusses the state of the art in probabilistic electric load forecasting. It differentiates between techniques and methodologies. With techniques they refer to a family of models, like multiple linear regression or artificial neural networks. On the other hand, methodologies consist of general frameworks that can be incorporated into any method, for example variable selection mechanisms. Also this paper stresses the need for some guidelines to standardize research in the field.

@@

Ziel et al.\ \cite{ziel2018probabilistic} provides detailed tables, groups research papers by the time dimension and objective and reports dataset, model and accuracy 

@@

David et al.\ \cite{david2016probabilistic} adopt a combination of autoregressive moving average (ARMA) and generalized autoregressive conditional heteroskedasticity (GARCH) in probabilistic forecasts of solar irradiance. Furthermore, they propose a recursive framework for parameter estimation.

@@

Van der Meer et al.\ \cite{van2018review} provides another thorough anlysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.

@@

Nowotarski et al.\ \cite{nowotarski2015computing} introduce a method for constructing prediction intervals (PI) and call it quantile regression averaging. Their idea is weighting a set of model predictions such that the pinball loss of the weighted model is minimized. The observed result is that QRA performed better compared to twelve individual models.

@@

It is also worth noting that older and simpler methods are still valuable (in combination with other models or on their own). These being less subject to overfitting than complex models.

@@

A major step forward in EF was the creation of the global energy forecasting competition (GEFCom) in 2012. Until then, no formal benchmarking process or data pool was established and new publications rarely reproduced the results from work done by others. Addressing these issues was the motivation behind the creation of GEFCom by the IEEE working group on energy forecasting. The EF field was positively affected by this competition. A number of ideas were tested on the same setting with only the best ones being published and it also contributed bridging the gap between industry practice and academic research. GEFCom 2012 had two tracks; the former about hierarchical load forecasting, the latter about wind power forecasting, see \cite{hong2014global} for a comprehensive review. 

@@

The focus of GEFCom 2014 was on probabilistic forecasting, Hong et al.\ \cite{hong2016probabilistic} discusses the problem tracks, the data and the winning methods.

@@

Xie et al.\ \cite{xie2016gefcom2014} propose a two stage approach; in the first stage they use multiple linear regression (MLR) to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried exponential smoothing (ESM), artificial neural networks (ANN) and autoregressive integrated moving average (ARIMA). 

@@

Gaillard et al.\ \cite{gaillard2016additive,gaillardasemi} combines quantile regression with generalized additive models \cite{hastie2017generalized}.

@@

The last GEFCom was held in 2017, its focus was providing probabilistic load forecasts, see \cite{hong2019global}. The GEFCom competition has also inspired the organization of other competitions such as the RWEnpower competition in the UK, the RTE competition in France, the Tokyo electric power company competition in Japan and the BigDEAL forecasting competitions.

@@

The EF field is characterized by heterogeneity in its forecasting techniques; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.

>>>>

%Electricity Forecasting 
%intro
During the past 25 years %because paper of 2014 says 15, so now it is 25
a wide range of new ideas have been proposed for electricity point forecasting and for probabilistic forecasting.
\\
The field benefitted greatly from the increase of computing power, the greater availability of data and the interest in data science.
As a consequence, the forecaster's toolbox has grown in size and complexity.
\\
Before delving into the literature review, we stress that at this point in time there is no superior method. Different solutions may outperform or underperform compared to other techniques depending on the problem settings. Thus, understanding the complexity, strengths and weaknesses of each method is crucial for fitting the right model to the right setting.
\\
Within this research community, the need for more homogeneity in the choice of the error valuation metrics emerged (Section \ref{metrics}), data quality and in the way of comparing model performances \cite{EPF_review}. As a solution, \cite{lago} proposes a checklist to aid evaluating the meaningfulness of new research. Throughout this thesis work, we will stick to the proposed principles and best practices peculiar of the EF field.
%literature review

@@

In terms of forecasting horizons, we can group EF into four major categories: very short term forecasting (VSTF), short term forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF). Consensus in the literature is to use one day, two weeks and three years respectively as cut off horizons \cite{hong_phd}; see Figure \ref{fig:time} for a visualisation. 

@@

EF literature distinguishes between point and probabilistic forecasts. Each of the two has its advantages and disadvantages. Point forecasts are easier to generate and less computationally intensive while probabilistic forecasts are more informative. Industry and research efforts have focused primarily on point forecasting. Nevertheless, interest in probabilistic forecasting has risen considerably over the last years due to renewable integration requirements, introduction of smart grids and increased market competitiveness.

@@

This section presents the results of the bibliometric analysis we performed on March $6^{th}$ 2024. This survey has been carried out by using the Scopus citation database. For details on the specific queries entered in Scopus, refer to Appendix \ref{src_code}.

@@

To get started, let us consider the evolution of the EF field over the years. This is visualised in Figure \ref{fig:epf_evolution}, with results grouped by category. Articles prior to the 2000 have been aggregated together due to their small number. Figure \ref{fig:epf_evolution} shows the trend of an increasing interest in EF.

@@

The next question is to compare the state of point-versus probabilistic forecasting, this is visualised in Figure \ref{fig:point_vs_prob}. What can be concluded is that probabilistic-is less developed than point forecasting. To our mind this is due to the complexity of probabilistic forecasts.

@@

EF is a heterogeneous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked. Figure \ref{fig:subject_area} depicts the EF publications by subject area. What can be concluded is that the bulk of publications come from engineering, computer science, mathematics and econometrics.

@@

  \caption{Most popular sources/outlets}

@@

Weron \cite{EPF_review} reviews the state of the art for electricity price forecasting. Besides analysing complexity of available solutions, strengths and weaknesses, it also stresses the need for objective comparative EPF studies. Specifically, it advocates for studies using similar datasets, using the same error evaluation metric and statistical testing model's outperformance.
Hong et al.\ \cite{prob_elf} discusses the state of the art in probabilistic electric load forecasting. It differentiates between techniques and methodologies. With techniques they refer to a family of models, like multiple linear regression or artificial neural networks. On the other hand, methodologies consist of general frameworks that can be incorporated into any method, for example variable selection mechanisms. Also this paper stresses the need for some guidelines to standardise research in the field.

@@

Ziel et al.\ \cite{ziel2018probabilistic} provides detailed tables, groups research papers by the time dimension and objective and reports datasets, models and accuracy 

@@

David et al.\ \cite{david2016probabilistic} adopts a combination of autoregressive moving average (ARMA) and generalised autoregressive conditional heteroskedasticity (GARCH) in probabilistic forecasts of solar irradiance. Furthermore, they propose a recursive framework for parameter estimation.

@@

Van der Meer et al.\ \cite{van2018review} provides another thorough analysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.

@@

Nowotarski et al.\ \cite{nowotarski2015computing} introduce a method for constructing prediction intervals (PI) and call it quantile regression averaging. Their idea is weighting a set of model predictions such that the pinball loss of the weighted model is minimised. The observed result is that QRA performed better compared to twelve individual models.

@@

It is also worth noting that older and simpler methods are still valuable (in combination with other models or on their own), these being less subject to overfitting than complex models.

@@

A major step forward in EF was the creation of the global energy forecasting competition (GEFCom) in 2012. Until then, no formal benchmarking process or data pool was established and new publications rarely reproduced the results from work done by others. Addressing these issues was the motivation behind the creation of GEFCom by the IEEE working group on energy forecasting. The EF field was positively affected by this competition. A number of ideas were tested on the same setting with only the best ones being published and it also contributed bridging the gap between industry practice and academic research. GEFCom2012 had two tracks; the former about hierarchical load forecasting, the latter about wind power forecasting, see \cite{hong2014global} for a comprehensive review. 

@@

The focus of GEFCom2014 was on probabilistic forecasting, Hong et al.\ \cite{hong2016probabilistic} discusses the problem tracks, the data and the winning methods.

@@

Xie et al.\ \cite{xie2016gefcom2014} proposes a two stage approach; in the first stage they use multiple linear regression (MLR) to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried exponential smoothing (ESM), artificial neural networks (ANN) and autoregressive integrated moving average (ARIMA). 

@@

Gaillard et al.\ \cite{gaillard2016additive,gaillardasemi} combines quantile regression with generalised additive models \cite{hastie2017generalized}.

@@

The last GEFCom was held in 2017, its focus was providing hierarchical probabilistic load forecasts, see \cite{hong2019global}. The GEFCom competition has also inspired the organisation of other competitions such as the RWEnpower competition in the UK, the RTE competition in France, the Tokyo electric power company competition in Japan and the BigDEAL forecasting competitions.

@@

The EF field is characterised by heterogeneity in its forecasting techniques; methods come from statisitics, mathematics, econometrics, electrical engineering and the artificial intelligence communities.

### 2024-06-03 thesis/input_files/metrics.tex
June 2nd 2024
final

\section{Root mean squared percentage error}\label{rmspe}
\begin{definition}
    The root mean squared percentage error is defined as
    $$
    \mathrm{RMSPE}(y,\hat{y})=100\cdot\sqrt{\frac{1}{h}\sum\limits_{k=1}^{h} \left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{|y_{n+k}|}\right)^2}
    $$
\end{definition}
Mean absolute percentage error and root mean squared percentage error may not be appropriate for series which have zero or very small values, for example, electricity demand at the household level. The result is a large score regardless of the absolute errors.
Scaled errors constitute a robust family of scores.
\section{Mean absolute scaled error}\label{mase}
\begin{definition}
    The mean absolute scaled error is defined as
    $$
    \mathrm{MASE}(y,\hat{y})=\frac{1}{h}\sum\limits_{k=1}^N\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}
    $$
\end{definition}
In the denominator we have the error of the naïve/persistence model. 
In this model, the current demand makes up the prediction for the next time step, that is $\hat{y}^{\mathrm{naive}}_{n+1}=y_{n}$.
\section{Root mean squared scaled error}\label{rmsse}
\begin{definition}
    The root mean squared scaled error is defined as
    $$
    \mathrm{RMSSE}(y,\hat{y})=\sqrt{\frac{1}{h}\sum\limits_{k=1}^N\left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}\right)^2}
    $$
\end{definition}

\section{Pinball}\label{pinball}

@@

\section{Winkler}

@@

    \mathrm{Winkler}(y_t, \hat{y}_t,q)=\begin{cases}

@@

Where $\delta$ is the prediction interval (PI) width, that is $\delta=u_t-l_t$, $u_t$ is the prediction interval upper threshold and $l_t$ is the prediction interval lower threshold at time step $t$. This score penalizes observations falling outside the prediction interval and rewards narrow prediction intervals.

@@

The probability integral transform (PIT) is a method to assess visually the quality of a probabilistic forecast. PIT is obtained by applying the predicted cumulative density function $\hat{F}$ to the data. If applying such CDF to the data results in a uniform distributed PIT, then $\hat{F}$ is a valid prediction. If not, $\hat{F}$ is not a good estimate of the CDF for the considered data. Figure \ref{fig:pit} provides an example, applying the true CDF results in a well calibrated PIT (left). Alternatively, applying a bad CDF results in either a overdispersed (middle) or underdispersed (right) PIT.

>>>>

\section{Root mean squared percentage error}\label{rmspe}
\begin{definition}
    The root mean squared percentage error is defined as
    $$
    \mathrm{RMSPE}(y,\hat{y})=100\cdot\sqrt{\frac{1}{h}\sum\limits_{k=1}^{h} \left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{|y_{n+k}|}\right)^2}
    $$
\end{definition}
Mean absolute percentage error and root mean squared percentage error may not be appropriate for series which have zero or very small values, for example, electricity demand at the household level. The result is a large score regardless of the absolute errors.
Scaled errors constitute a robust family of scores.
\section{Mean absolute scaled error}\label{mase}
\begin{definition}
    The mean absolute scaled error is defined as
    $$
    \mathrm{MASE}(y,\hat{y})=\frac{1}{h}\sum\limits_{k=1}^h\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}
    $$
\end{definition}
In the denominator we have the error of the naïve/persistence model. 
In this model, the current demand makes up the prediction for the next time step, that is $\hat{y}^{\mathrm{naive}}_{n+1}=y_{n}$.
\section{Root mean squared scaled error}\label{rmsse}
\begin{definition}
    The root mean squared scaled error is defined as
    $$
    \mathrm{RMSSE}(y,\hat{y})=\sqrt{\frac{1}{h}\sum\limits_{k=1}^h\left(\frac{|y_{n+k}-\hat{y}_{n+k}|}{\frac{1}{h-1}\sum\limits_{k=2}^{h}|y_{k}-y_{k-1}|}\right)^2}
    $$
\end{definition}

\section{Pinball score}\label{pinball}

@@

\section{Winkler score}

@@

    \mathrm{Winkler}(y_t,  u_{t},l_{t},\alpha)=\begin{cases}

@@

Where $\delta$ is the prediction interval (PI) width, that is $\delta=u_t-l_t$, $\alpha$ is (1-coverage level), $u_t$ is the prediction interval upper threshold and $l_t$ is the prediction interval lower threshold at time step $t$. This score penalises observations falling outside the prediction interval and rewards narrow prediction intervals.

@@

The probability integral transform (PIT) is a method to assess visually the quality of a probabilistic forecast. PIT is obtained by applying the predicted cumulative density function $\hat{F}$ to the data. If applying such CDF to the data results in a uniform distributed PIT, then $\hat{F}$ is a valid prediction. If not, $\hat{F}$ is not a good estimate of the CDF for the considered data. Figure \ref{fig:pit} provides an example, applying the true CDF results in a well calibrated PIT (left). Alternatively, applying a bad CDF results in either an overdispersed (middle) or underdispersed (right) PIT.
